State and Input Dependent Disturbances in Min-Max Model Predictive Control

### 2015-11-16 draft.tex
more edits to introduction

A major challenge in RMPC is to guarantee feasibility of the control problem for all possible uncertainties in 
presence of state constraints.
%
This can be achieved by introducing additional constraints into the problem that force the state at the end of the horizon to lie in a set that is feasible and invariant for all possible realisations of uncertainty under a given feedback law; the largest such set, 
the so-called \emph{maximal robust positively invariant} (MRPI) set, is often used for this purpose.
%
The concept of MRPI sets and their analytical properties were introduced in the literature (e.g.~\cite{Bertsekas:1971,Glover:1971}) shortly after robust model predictive control was originally proposed. However algorithms for numerically 
computing MRPI sets were not developed until more than two decades later~\cite{Blanchini:1994,DeSantis:1994,Kolmanovsky:1998};
earlier algorithms either failed to ensure finite determinability
of the MRPI set (see e.g.~\cite{Blanchini:1990}), 
or were not guaranteed to produce the maximal RPI set (e.g.~\cite{Blanchini:1991}).
%

This paper considers the use of parametrised sets for model uncertainties, which were recently introduced in~\cite{Schaich:CDC:2015,Schaich:NMPC:2015}. As parameters we consider the state and the control input; this allows linearisation errors and multiplicative uncertainties to be modelled as additive disturbances.
%
The paper's contribution is to develop
%extends~\cite{Schaich:CDC:2015} by developing 
the analytical and computational tools that are needed to characterise and determine invariant and controllable sets for systems with parametrised uncertainty sets, and to implement them in a closed loop RMPC strategy. To this end, we provide necessary and sufficient conditions for the property, introduced in~\cite{Schaich:CDC:2015}, of \emph{parametric convexity}, which is essential in order to obtain convex MRPI and controllable sets, and we derive a bound on the number of iterations needed to compute the MRPI set. The current paper also develops and extends the min-max RMPC approach introduced in~\cite{Schaich:NMPC:2015} by considering the case of degenerate constraints.

To solve RMPC problems a few different methods are available. For the case of unknown additive disturbances, 
a scenario-based approach is proposed in~\cite{Scokaert:1998} for the online solution of min-max RMPC problems. However, the computational complexity of this approach grows exponentially with the length of the horizon. 
%
A parametrisation of predicted feedback laws proposed in~\cite{Rakovic:2012} reduces the RMPC problem to a single linear 
program to be solved online, but although this is a closed loop strategy the parametrisation is in general 
suboptimal. 
%
Explicit RMPC algorithms (e.g.~\cite{Bemporad:2003,Diehl:2004}) optimise the worst case predicted performance for each admissible model state offline and
use results from multi-parametric programming to parametrise the optimal control law as a piecewise affine function of the
state.
%
However, this requires extensive offline computation and becomes impractical for a high-dimensional state space.
%
In this paper we use an active-set approach based on that presented in~\cite{Buerger:ACC,Buerger:IJRNC,Schaich:NMPC:2015}.
In a similar manner to explicit RMPC algorithms, we use multi-parametric programming to obtain solutions efficiently
for a given active-set, however, we avoid an exploration of the entire feasible set of states but rather solve online for the current state only.
%
Our approach uses a line search to update the set of active constraints, thus generating
the solution of the RMPC problem online in an efficient manner.

The paper is structured as follows, in section~\ref{sec:problem:statement} we present the RMPC problem under consideration
and define the constraint sets for the model uncertainty in terms of state and input dependent disturbance sets.
%
The properties of parametrised sets are then considered in section~\ref{sec:p:convex:sets}, in particular we discuss
parametric convexity and its characterisation in section~\ref{ssec:properties:of:p:convex:maps}, and its use in the computation of MRPI sets, which is presented in section~\ref{ssec:MRPI:sets}.
%
To illustrate how state and input dependent disturbances can be used to account for nonlinearities and 
calculate the MRPI set we present an example in section~\ref{sec:example:I}.
%
The solver we present here relies entirely on a recursive use of multi-parametric programming techniques, 
which are presented in section~\ref{ssec:recursive:mpQP} for a given active-set, the active set is then
updated using a line search which is described in subsection~\ref{ssec:line:search}. In order to account for the 
degeneracy that can occur in the optimisation, we consider the case of degenerate problems in subsection~\ref{ssec:degeneracy}.
%
In section~\ref{sec:solving:min:max:programs} we translate the general formulation of section~\ref{sec:recursive:mpQP:via:line:search}
to the problem presented in section~\ref{sec:problem:statement}, and section~\ref{sec:complexity}
deals with the complexity of the proposed algorithm.
%
We revisit the example system presented in section~\ref{sec:example:I} to illustrate the proposed RMPC
algorithm in section~\ref{sec:example:II}.
%
Section~\ref{sec:conclusion} provides brief conclusions.

@@

for elaborate discussions) can be solved efficiently with an active-set solver (see e.g.~\cite{Fletcher:2000}).

@@

We want to use an active-set solver to obtain the solution for $\phi=\phi^\ast$.

@@

A standard non-degeneracy assumption for active-set solvers is that only one constraint is affected at each line-search
iteration, this means that when constraints become degenerate the null-space~$\ker C_{\mathcal A}$ is
one-dimensional, furthermore no additional constraints may be activated since~$\phi$ is already fixed by
the constraints.
%
This implies that~\eqref{mpQP:linesearch:degenerate:constraints:beta} is a scalar linear program, since
$O_{\hat\beta}$ parametrises the null-space of $\ker C_{\mathcal A}$.
%
We conclude that no more than two constraints are candidates to be deactivated in order to remove the degeneracy
of the constraints, namely the ones corresponding to the maximal and minimal $\hat\beta=
-\frac{O_\phi\phi+o}{O_{\hat\beta}}$.
%
After deactivating the constraint corresponding to the maximiser of~\eqref{mpQP:linesearch:degenerate:constraints:beta},
the line-search can be continued without redundancy.

@@

form a known active-set~${\bar{\mathcal A}} = \{\mathcal A_N,\hat {\mathcal A}_N,\dots,\hat{\mathcal A}_1\}$

>>>>

This paper considers the use of parametrised sets to bound model uncertainty. By allowing the bounds on model uncertainty to depend explicitly on the model state and control input we obtain a generalised framework for RMPC in which linearisation errors and multiplicative uncertainties can be modelled as additive disturbances.
%
General parametrised disturbance bounding sets were used in~\cite{Rakovic03,Rakovic06}; however, here we focus exclusively on problem formulations that lead to convex-concave min-max control problems.
% 
The paper's main contributions are to develop
%extends~\cite{Schaich:CDC:2015} by developing 
the analytical and computational tools that are needed to characterise and determine convex invariant and controllable sets for systems with parametrised uncertainty sets, and to employ these sets in a closed loop RMPC strategy. To this end, we provide necessary and sufficient conditions for the property, introduced in~\cite{Schaich:CDC:2015}, of \emph{parametric convexity}. We show that this condition is both necessary and sufficient to ensure that the corresponding robustly invariant and controllable sets are convex.

A major challenge in RMPC is to guarantee feasibility of the control problem for all possible uncertainties in 
presence of state constraints.
%
This can be achieved by introducing additional constraints into the problem that force the state at the end of the horizon to lie in a set that is feasible and invariant for all possible realisations of uncertainty under a given feedback law; the largest such set, 
the so-called \emph{maximal robust positively invariant} (MRPI) set, is often used for this purpose.
%
The concept of MRPI sets and their analytical properties were introduced in the literature (e.g.~\cite{Bertsekas:1971,Glover:1971}) shortly after robust model predictive control was originally proposed. However algorithms for numerically 
computing MRPI sets were not developed until more than two decades later~\cite{Blanchini:1994,DeSantis:1994,Kolmanovsky:1998}.
% earlier algorithms either failed to ensure finite determinability
% of the MRPI set (see e.g.~\cite{Blanchini:1990}), 
% or were not guaranteed to produce the maximal RPI set (e.g.~\cite{Blanchini:1991}).
%
This paper builds on the work of~\cite{Kolmanovsky:1998,Rakovic06} by proposing an iterative procedure for computing the MRPI set for the case of parametrically convex state and input dependent disturbance sets. We show that, under mild assumptions, this procedure determines the MRPI set in a finite number of iterations, and we provide a bound on the number of iterations required.

The current paper also develops an online implementation of closed loop min-max RMPC that allows the bounds on model uncertainty to depend on the model state and control input.
%
Several methods are available to solve RMPC problems with fixed disturbance bounds. For the case of unknown additive disturbances, 
a scenario-based approach is proposed in~\cite{Scokaert:1998} for the online solution of min-max RMPC problems. However, the computational complexity of this approach grows exponentially with the length of the horizon. 
%
A parametrisation of predicted feedback laws proposed in~\cite{Rakovic:2012} reduces the RMPC problem to a single linear 
program to be solved online, but although this is a closed loop strategy, the parametrisation is in general 
suboptimal. 
%
The Explicit RMPC algorithms~\cite{Bemporad:2003,Diehl:2004} optimise the worst case predicted performance for each admissible model state offline and
use results from multi-parametric programming to parametrise the optimal control law as a piecewise affine function of the
state.
%
However, this requires extensive offline computation and becomes impractical for a high-dimensional state space.
%
In this paper we use an active set approach based on that presented in~\cite{Buerger:ACC,Buerger:IJRNC} for the case of additive disturbances confined to a fixed set and we extend the approach for parametrised sets that was introduced in~\cite{Schaich:NMPC:2015} by considering the case of degenerate constraints.
%
In a similar manner to explicit RMPC algorithms, we use multi-parametric programming to obtain solutions efficiently
for a given active set, however, we avoid an exploration of the entire feasible set of states but rather solve online for the current state only.
%
Our approach uses a line search to update the set of active constraints, thus generating
the solution of the RMPC problem online in an efficient manner.

The paper is structured as follows, in section~\ref{sec:problem:statement} we present the RMPC problem under consideration
and define the constraint sets for the model uncertainty in terms of state and input dependent disturbance sets.
%
The properties of parametrised sets are then considered in section~\ref{sec:p:convex:sets}, in particular we discuss
parametric convexity and its characterisation in section~\ref{ssec:properties:of:p:convex:maps}, and its use in the computation of MRPI sets, which is presented in section~\ref{ssec:MRPI:sets}.
%
To illustrate how state and input dependent disturbances can be used to account for nonlinearities and 
calculate the MRPI set we present an example in section~\ref{sec:example:I}.
%
The solver we present here relies entirely on a recursive use of multi-parametric programming techniques, 
which are presented in section~\ref{ssec:recursive:mpQP} for a given active set, the active set is then
updated using a line search which is described in subsection~\ref{ssec:line:search}. In order to account for the 
degeneracy that can occur in the optimisation, we consider the case of degenerate problems in subsection~\ref{ssec:degeneracy}.
%
In section~\ref{sec:solving:min:max:programs} we translate the general formulation of section~\ref{sec:recursive:mpQP:via:line:search}
to the problem presented in section~\ref{sec:problem:statement}, and section~\ref{sec:complexity}
deals with the complexity of the proposed algorithm.
%
We revisit the example system presented in section~\ref{sec:example:I} to illustrate the proposed RMPC
algorithm in section~\ref{sec:example:II}.
%
Section~\ref{sec:conclusion} provides brief conclusions.

@@

for elaborate discussions) can be solved efficiently with an active set solver (see e.g.~\cite{Fletcher:2000}).

@@

We want to use an active set solver to obtain the solution for $\phi=\phi^\ast$.

@@

A standard non-degeneracy assumption for active set solvers is that only one constraint is affected at each line-search
iteration, this means that when constraints become degenerate the null-space~$\ker C_{\mathcal A}$ is
one-dimensional, furthermore no additional constraints may be activated since~$\phi$ is already fixed by
the constraints.
%
This implies that~\eqref{mpQP:linesearch:degenerate:constraints:beta} is a scalar linear program, since
$O_{\hat\beta}$ parametrises the null-space of $\ker C_{\mathcal A}$.
%
We conclude that no more than two constraints are candidates to be deactivated in order to remove the degeneracy
of the constraints, namely the ones corresponding to the maximal and minimal $\hat\beta=
-\frac{O_\phi\phi+o}{O_{\hat\beta}}$.
%
After deactivating the constraint corresponding to the maximiser of~\eqref{mpQP:linesearch:degenerate:constraints:beta},
the line-search can be continued without redundancy.

@@

form a known active set~${\bar{\mathcal A}} = \{\mathcal A_N,\hat {\mathcal A}_N,\dots,\hat{\mathcal A}_1\}$
